Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-04-13 papers/kmeans/kmeans.tex
Addressed issues raised by the referee

\begin{abstract}
% Context:
Chemical tagging promises to use detailed abundance measurements to
identify spatially separated stars that were in fact born together
(in the same molecular cloud), long ago.
This idea has not previously yielded scientific successes, probably because of the
noise and incompleteness in chemical-abundance measurements.
However, we have succeeded in substantially improving spectroscopic
measurements with \thecannon, which has delivered $15$ individual
abundances for $\sim 10^5$ stars observed as part of the
\apogee\ spectroscopic survey, with precisions around 0.04~dex.
% Aims:
We test the chemical-tagging hypothesis by looking at clusters in abundance space
and confirming that they are clustered in phase space.
% Methods:
We identify (by the k-means algorithm) overdensities of stars
in the 15-dimensional chemical-abundance space delivered by \thecannon,
and plot the associated stars in phase space.
We use \emph{only} abundance-space information (no positional information) to identify stellar groups.
% Results:
We find that clusters in abundance space are indeed clusters in phase space.
We recover some known phase-space clusters and find other interesting structures.
This confirms the chemical-tagging hypothesis and verifies the precision of the abundance measurements delivered by \thecannon.
This is the first-ever project to identify phase-space structures by blind search purely in abundance space;
the prospects for future data sets are very good.
\end{abstract}

@@

Chemical tagging holds the promise of revealing not just the
star-formation history of the Galaxy, but also the accretion history
(as things that fall in are expected to be chemically distinct from
those that form in the parent body; for example, \citealt{font}) and
stellar-orbit diffusion process like radial mixing and radial
migration (for example, \citealt{roskar, quillen}).  After
stars are born---or after a star cluster is accreted and
disrupted---associations or groups will disperse, through
two-body processes, interactions with resonances, or tides from the whole Galaxy.

@@

It is important to realize that analysis techniques based on \foreign{ab-initio}
physical models are inaccurate, yet they too strive to improve precision
in the face of knowingly inaccurate models.  Incomplete atomic data, 
simplifications of photospheric structure, assumptions about convective
motion, and inconsistencies resulting from positing local thermal
equilibrium all contribute to produce inaccurate abundances.
\thecannon\ stands out because it demonstrably improves the precision on
chemical abundances, whereas the accuracy of those labels is limited only
by the training (reference) set employed.

@@

Six two-dimensional projections of the abundance data, plus some other
data quantities, are shown in \figurename~\ref{fig:all}.  Throughout
this work we show two-dimensional abundance projections for only a few
elements. The entire sequence of 15-choose-2 combinations is too
immense to visualize. The projections shown are ones which
have been extensively used in abundance studies of globular clusters
and satellite systems. These are typically light element abundances,
and our projections include C--N, Na--O, and Mg--Al. Globular clusters
famously demonstrate correlations in these elements, thereby making
them suitable projections for us to show in verifying and examining
any substructure identified by k-means (or any other clustering algorithm).

@@

Nothing about this algorithm or our choices are particularly
tuned or optimized; this is in no sense the
algorithm or the method that reveals the best structures.
Indeed, it is a feature of this work that even the simplest, most
generic clustering algorithm returns interesting
structures (as shown below).
The prospects for Milky Way science only improve as the clustering
algorithm improves.

@@

@@

@@

@@

>>>>

\begin{abstract}
% Context:
Chemical tagging promises to use detailed abundance measurements to
identify spatially separated stars that were in fact born together
(in the same molecular cloud), long ago.
This idea has not previously yielded scientific successes, probably because of the
noise and incompleteness in chemical-abundance measurements.
However, we have succeeded in substantially improving spectroscopic
measurements with \thecannon, which has delivered $15$ individual
abundances for $\sim 10^5$ stars observed as part of the
\apogee\ spectroscopic survey, with precisions around 0.04~dex.
% Aims:
We test the chemical-tagging hypothesis by looking at clusters in abundance space
and confirming that they are clustered in phase space.
% Methods:
We identify (by the k-means algorithm) overdensities of stars
in the 15-dimensional chemical-abundance space delivered by \thecannon,
and plot the associated stars in phase space.
We use \emph{only} abundance-space information (no positional information) to identify stellar groups.
% Results:
We find that clusters in abundance space are indeed clusters in phase space.
We recover some known phase-space clusters and find other interesting structures.
This confirms the chemical-tagging hypothesis and verifies the precision of the abundance measurements delivered by \thecannon.
This is the first-ever project to identify phase-space structures at survey-scale by blind search purely in abundance space;
the prospects for future data sets are very good.
\end{abstract}

@@

Chemical tagging holds the promise of revealing not just the
star-formation history of the Galaxy, but also the accretion history
(as things that fall in are expected to be chemically distinct from
those that form in the parent body; for example, 
\citealt{font,De_Silva_2007,eggen_1970,bubar_king_2010}) and
stellar-orbit diffusion process like radial mixing and radial
migration (for example, \citealt{roskar, quillen}).  After
stars are born---or after a star cluster is accreted and
disrupted---associations or groups will disperse, through
two-body processes, interactions with resonances, or tides from the whole Galaxy.

@@

It is important to realize that analysis techniques based on \foreign{ab-initio}
physical models are inaccurate, yet they too strive to improve precision
in the face of knowingly inaccurate models.  Incomplete atomic data, 
simplifications of photospheric structure, assumptions about convective
motion, and inconsistencies resulting from positing local thermal
equilibrium all contribute to produce inaccurate abundances.
\thecannon\ stands out because it demonstrably improves the precision on
chemical abundances, whereas the accuracy of those labels is limited only
by the training (reference) set employed.  At the same time, it is crucial
to be cognisant of the constraints in the training set: the
results of \thecannon\ will be limited by the quality of the training 
set labels. While chemical tagging does not necessarily require
accurate abundance labels (precision is paramount), comparing abundance
labels to models of Galactic chemical enrichment requires a firm level
of belief in the label accuracy. Particularly for the abundance labels
of stars in the training set.

@@

Six two-dimensional projections of the abundance data, plus some other
data quantities, are shown in \figurename~\ref{fig:all}.  Throughout
this work we show two-dimensional abundance projections for only a few
elements. The entire sequence of 15-choose-2 combinations is too
immense to visualize. The projections shown are ones which
have been extensively used in abundance studies of globular clusters
and satellite systems. These are typically light element abundances,
and our projections include C--N, Na--O, and Mg--Al. Globular clusters
famously demonstrate correlations in these elements \citep[e.g., see][and references therein]{Norris_1995,Carretta_2009}, thereby making
them suitable projections for us to show in verifying and examining
any substructure identified by k-means (or any other clustering algorithm).

@@

Nothing about this algorithm or our choices are particularly
tuned or optimized; this is in no sense the
algorithm or the method that reveals the best structures.
We chose k-means as a simple and tractable approach to identifying
clusters in high dimensional label space. In practice more complex
clustering algorithms may perform better than k-means, or even
tuning k-means by heuristically setting $K$ is likely to improve
upon the results here. This \documentname\ serves as a proof of concept.
Indeed, it is a feature of this work that even the simplest, most
generic clustering algorithm returns interesting
structures (as shown below). 
The prospects for Milky Way science only improve as the clustering
algorithm improves.

@@

\bibitem[Bubar \& King(2010)]{bubar_king_2010} Bubar, E.~J., \& King, J.~R.\ 2010, \aj, 140, 293 
\bibitem[Carretta et al.(2009)]{Carretta_2009} Carretta, E., Bragaglia, A., Gratton, R.~G., et al.\ 2009, \aap, 505, 117 

@@

\bibitem[De Silva et al.(2007)]{De_Silva_2007} De Silva, G.~M., Freeman, K.~C., Bland-Hawthorn, J., Asplund, M., \& Bessell, M.~S.\ 2007, \aj, 133, 694

@@

\bibitem[Eggen(1970)]{eggen_1970} Eggen, O.~J.\ 1970, \pasp, 82, 99 

@@

\bibitem[Norris \& Da Costa(1995)]{Norris_1995} Norris, J.~E., \& Da Costa, G.~S.\ 1995, \apjl, 441, L81 
